Brother, Can You Spare a Manufacturing Job? How Voters React to Deindustrialization

### 2021-12-10 Final Draft/Final Draft.Rmd
more writing

---
title: "Brother, Can You Spare a Manufacturing Job?  How Voters React to Deindustrialization"
short: ""
journal: "" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 1
issue: 1
author:
  - name: Catherine Darin
    firstname: Catherine
    surname: Darin
    email: cdarin@hks.harvard.edu
    affiliation: Harvard University
  - name: Zagreb Mukerjee
    firstname: Zagreb
    surname: Mukerjee
    email: zagrebmukerjee@fas.harvard.edu
    affiliation: Harvard University
acknowledgements: 
abstract: |
  What led to Donald Trump's surprising 2016 election victory? This paper examines the potential contribution of deindustrialization. Counties differ by exposure to manufacturing, which we use in an instrumental-variable approach to identify the effect of manufacturing job loss on change in Democratic vote share. We find that Democratic vote share falls in counties with high manufacturing job loss since 2004. Disaggregating this result by race shows that the Democratic vote share fell where layoffs affected white populations, and rose where layoffs affected nonwhites. This suggests a racial component to how voters process economic hardship.
output: rticles::aea_article
---

```{r}
library(tidyverse)
library(stargazer)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)
library(readxl)
library(scales)
library(tigris)

#knitr::opts_chunk$set(fig.pos = "!H", out.extra = "")
#knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

@@

We replicate Baccini and Wemouth (2021), which analyzes the effect of gross manufacturing job losses from 2012 to 2015 on the change in Democratic vote share in the 2016 presidential election.

@@

Following \cite{BW}, our unit of analysis is the county. This allows us to capture not only the direct effects of deindustrialization on laid-off manufacturing workers, but also on the local economy. The primary outcome variable is the county's change in Democratic vote share between 2012 and 2016. We obtain this from XXXXXXXXX

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the destruction of manufacturing jobs, as a share of the total employment in each county.

>>>>

---
title: "Brother, Can You Spare a Manufacturing Job?  How Voters React to Deindustrialization"
short: "Brother, Can You Spare a Manufacturing Job? "
journal: "" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 1
issue: 1
author:
  - name: Catherine Darin
    firstname: Catherine
    surname: Darin
    email: cdarin@hks.harvard.edu
    affiliation: Harvard University
  - name: Zagreb Mukerjee
    firstname: Zagreb
    surname: Mukerjee
    email: zagrebmukerjee@fas.harvard.edu
    affiliation: Harvard University
acknowledgements: 
abstract: |
  What led to Donald Trump's surprising 2016 election victory? This paper examines the potential contribution of deindustrialization. Counties differ by exposure to manufacturing, which we use in an instrumental-variable approach to identify the effect of manufacturing job loss on change in Democratic vote share. We find that Democratic vote share falls in counties with high manufacturing job loss since 2004. Disaggregating this result by race shows that the Democratic vote share fell where layoffs affected white populations, and rose where layoffs affected nonwhites. This suggests a racial component to how voters process economic hardship.
output: rticles::aea_article
---

@@

We replicate \cite{BW21}, which analyzes the effect of gross manufacturing job losses from 2012 to 2015 on the change in Democratic vote share in the 2016 presidential election.

@@

Following \cite{BW21}, our unit of analysis is the county. This allows us to capture not only the direct effects of deindustrialization on laid-off manufacturing workers, but also on the local economy. The primary outcome variable is the county's change in Democratic vote share between 2012 and 2016. We obtain this from XXXXXXXXX

We want to measure the causal relationship between deindustrialization and the change in vote share. As a measure of deindustrialization, we use the loss of manufacturing jobs as a share of the total beginning-of-period employment in each county. Suppose at the beginning of our period there were $2000$ manufacturing workers in a county, and $8000$ non-manufacturing workers. At the end there are only $1500$ manufacturing workers. This is a loss of $500/10000 = 5\%.$

In this choice we diverge from \cite{BW21}, which uses **gross** manufacturing job losses. For example, if a county lost $450$ manufacturing jobs from 2012-2015 and also gained $400$ manufacturing jobs over the same period, this would be $450$ **gross** job losses, but only $50$ **net** job losses. The **gross** measure captures several dynamics unrelated to deindustrialization - for example, seasonal unemployment in a food-manufacturing region, or workers moving between jobs. Thus we believe the **net** job losses over a period are the more accurate measure.

Data on job gains and losses are obtained, using an API, from the Census Bureau's Quarterly Workforce Indicators (\cite{QWI}), which contains (among other things) information about employment by industry and county. These statistics are further disaggregated by race and ethnicity. The Census Bureau obtains this data from a combination of sources, such as administrative tax data and the U.S. Census.

XXXXXXXXXXXXXXXX sources of other data

\subsection{Methods}

We conduct two related tests of our hypothesis. Our first test involves an instrumental-variable approach. A regression of change in Democratic vote share on manufacturing job losses risks endogeneity, in case counties with manufacturing job losses were otherwise predisposed to turn towards Trump (a singular candidate, after all). To mitigate these risks, \cite{BW21} uses a Bartik instrument, which we adapt. This instrument essentially uses the cross-county distribution of manufacturing employment as a source of exogenous variation.

$$
\begin{aligned}
b_{j,c} &=& \frac{\text{Manufacturing Employment}_{j,c} \text{ at } t_0}{\text{Total Employment}_c \text{ at }t_0}  \\
&&  * \frac{\text{National Manufacturing Job Change}_{j,c} }{\text{Total National Employment at }t_0} \\
\end{aligned}
$$

### 2021-12-10 Final Draft/Final-Draft.tex
more writing

>>>>

\section{Data and Methods}

\subsection{Data}

Following \cite{BW21}, our unit of analysis is the county. This allows
us to capture not only the direct effects of deindustrialization on
laid-off manufacturing workers, but also on the local economy. The
primary outcome variable is the county's change in Democratic vote share
between 2012 and 2016. We obtain this from XXXXXXXXX

We want to measure the causal relationship between deindustrialization
and the change in vote share. As a measure of deindustrialization, we
use the loss of manufacturing jobs as a share of the total
beginning-of-period employment in each county. Suppose at the beginning
of our period there were \(2000\) manufacturing workers in a county, and
\(8000\) non-manufacturing workers. At the end there are only \(1500\)
manufacturing workers. This is a loss of \(500/10000 = 5\%.\)

In this choice we diverge from \cite{BW21}, which uses \textbf{gross}
manufacturing job losses. For example, if a county lost \(450\)
manufacturing jobs from 2012-2015 and also gained \(400\) manufacturing
jobs over the same period, this would be \(450\) \textbf{gross} job
losses, but only \(50\) \textbf{net} job losses. The \textbf{gross}
measure captures several dynamics unrelated to deindustrialization - for
example, seasonal unemployment in a food-manufacturing region, or
workers moving between jobs. Thus we believe the \textbf{net} job losses
over a period are the more accurate measure.

Data on job gains and losses are obtained, using an API, from the Census
Bureau's Quarterly Workforce Indicators (\cite{QWI}), which contains
(among other things) information about employment by industry and
county. These statistics are further disaggregated by race and
ethnicity. The Census Bureau obtains this data from a combination of
sources, such as administrative tax data and the U.S. Census.

XXXXXXXXXXXXXXXX sources of other data

\subsection{Methods}

We conduct two related tests of our hypothesis. Our first test involves
an instrumental-variable approach. A regression of change in Democratic
vote share on manufacturing job losses risks endogeneity, in case
counties with manufacturing job losses were otherwise predisposed to
turn towards Trump (a singular candidate, after all). To mitigate these
risks, \cite{BW21} uses a Bartik instrument, which we adapt. This
instrument essentially uses the cross-county distribution of
manufacturing employment as a source of exogenous variation.

\[
\begin{aligned}
b_{j,c} &=& \frac{\text{Manufacturing Employment}_{j,c} \text{ at } t_0}{\text{Total Employment}_c \text{ at }t_0}  \\
&&  * \frac{\text{National Manufacturing Job Change}_{j,c} }{\text{Total National Employment at }t_0} \\
\end{aligned}
\]
